---
title: "Stats 20, F23 -- Homework 3"
author: "Michelle Navarrete Vega -- 805676988"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1a}
commute_times <- c(14, 12, 20, 19, 15, 20, 28, 20, 20, 18)

commute_sd <- sd(commute_times)
commute_mean <- mean(commute_times)

commute_sd

commute_times[(abs(commute_times-commute_mean))>(commute_sd)]

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->
The commute times were 14, 12, and 28 minutes. The standard deviation is 4.40202. The difference between the mean and the commute time is greater than 4.40202 for these days.

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1b}

commute_times[(abs(commute_times-commute_mean))<(commute_sd)]

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->
With standard deviation of the mean ==> mean - SD to mean + SD
Thus, within SD, then the distance between the time and the mean must be less than one SD.

### (c)
<!-- USE THE CODE CHUNK BELOW -->

Using the same logical index from (a), what proportion of days did Andy have a commute time that was
within one standard deviation of the mean commute time?
Hint: Can arithmetic operators/functions for numeric vectors work for logical vectors? What do sum() and mean() compute for logical vectors?

```{r Question 1c}

commute_times[(abs(commute_times-commute_mean))>(commute_sd)]
within_SD_logical <- (abs(commute_times-commute_mean))>(commute_sd)

sum(within_SD_logical)

prop_withinSD <- sum(within_SD_logical)/length(commute_times)
prop_withinSD
```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->
I made a logical vector using the logical index. Then by adding up the vector, the sum function counts how many 'TRUE' values the vector has. Thus, we get the proportion by dividing that sum by the length. 

## Question 2

```{r Question 2}
NA & TRUE
NA & FALSE
NA | TRUE
NA | FALSE
```
<!-- ENTER YOUR EXPLANATION BELOW, FEEL FREE TO EDIT THE ABOVE CODE CHUNK -->
Since NA stands for not available and used when a value is missing, when doing a & operation, we must look at both sides of the operation. 

For NA & TRUE, we cannot say FALSE because we do not know what the left side is since it is NA, and thus we get NA because if NA = false then the entire operation is false, but if it is true, since the right side is also true, the operation will be true. Thus, whatever NA is, that is what the result will be for this specific operation.

For NA & FALSE, it does not matter what NA is because we already know the entire thing will be false since the right side is false. Thus, there is no need to see what NA is. 

For NA | TRUE, we do not know what NA is, however this is an 'or' operation and since one side is TRUE, the entire thing is true, regardless of what NA is. 

For NA | FALSE, we do not know what NA is, and since this is an 'or' operation, if NA is true, then the statement will be true. However, if NA is false, since the right side is also false, the entire thing will be false. Here, the result is dependent on what NA is which is why it is the final result. 


## Question 3

<!-- PUT YOUR get_minimum_coins() FROM HW 1 HERE -->
```{r Question 3 background} 
  get_minimum_coins <- function(cents){
    
  num_of_quarters <- 0
  num_of_dimes <- 0
  num_of_nickels <- 0
  num_of_pennies <- 0
  
  num_of_quarters <- cents %/% 25
  
  num_of_dimes <- (cents %% 25) %/% (10)
  
  num_of_nickels <- ((cents %% 25) %% 10) %/% (5)
  
  remaining <- ((cents %% 25) %% 10)
  
  num_of_pennies <- remaining %% 5

  total <- num_of_quarters + num_of_dimes + num_of_nickels + num_of_pennies
  total
}
```

### (a)
<!-- ENTER YOUR ANSWER BELOW -->

<!-- USE THE CODE CHUNK BELOW IF YOU'D LIKE -->
```{r Question 3a}
minimum_vector <- c(5, 78, 6)
get_minimum_coins(5)
get_minimum_coins(78)
get_minimum_coins(6)
get_minimum_coins(minimum_vector)
```
It is vectorized since the operations inside the function are vectorized as well. 

### (b)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST COMMANDS; NO NEED TO STORE OBJECTs -->
```{r Question 3b}
num_of_cents <- 1:100
num_of_cents[get_minimum_coins(num_of_cents) == max(get_minimum_coins(num_of_cents))]
  
```

## Question 4

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 4a}

my_min <- function(vec, na.rm = FALSE){
  vec_min <- 0

if(sum(is.na(vec)) > 1){
    vec_min <- NA
    vec_min
    } else{
          vec_min <- vec[1]
  }
  
  
  vec_min <- vec[1]
  
  for(i in 1:length(vec)){
     if(!is.na(vec[i]) & (vec[i] < vec_min)){
      vec_min <- vec[i]}
  }
    

  vec_min
  
}




```

###
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 4b}

my_min(c(4, 1, 0, 2, -3, -5, -4)) == min(c(4, 1, 0, 2, -3, -5, -4))
my_min(c("bears", "beets", "Battlestar Galactica")) == min(c("bears", "beets", "Battlestar Galactica"))
my_min(7) == min(7)

my_min(c("Pawnee", "rules", "Eagleton", NA), na.rm = TRUE) == min(c("Pawnee", "rules", "Eagleton", NA), na.rm = TRUE)

my_min(c("Pawnee", "rules", "Eagleton", NA), na.rm = FALSE) == min(c("Pawnee", "rules", "Eagleton", NA), na.rm = FALSE)

my_min(NA, na.rm = FALSE) == min(NA, na.rm = FALSE)
my_min(NA, na.rm = TRUE) == min(NA, na.rm = TRUE)
min(NA, na.rm = TRUE)
my_min(NA, na.rm = TRUE)
```


## Question 5

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5a}
fib1 <- 1
fib2 <- 1
full_fib <- c(fib1, fib2)
while (fib1 + fib2 < 500) {
fib2 <- fib1 + fib2
full_fib <- c(full_fib, fib2)
fib1 <- fib2-fib1

}
full_fib

```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5b}
full_fib <- c(1, 1)
while (fib1+fib2 < 500) {
full_fib <- c(full_fib, full_fib)
full_fib <- full_fib*1/2
}
full_fib
```

### (c)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5c}
i <- 3
full_fib <- c(1, 1)
while ((full_fib[i-1]+full_fib[i-2]) < 500) {
      full_fib <- c(full_fib, (full_fib[i-1] + full_fib[i-2]))
      i <- i + 1
  }

full_fib

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL). ONE SENTENCE OR 'MY ANSWER IS __' IS FINE. -->

## Question 6
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6}
##jerry <- 2:8 *(5 %% 3^(-2:7 > 2))
##jerry
3^(-2:7 > 2)
(2:8 *5) %% 3^-2:(7 > 2)

5%%(1/3)
```

## Question 7

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNmENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7a}
my_ifelse <- function(test, yes, no){
   ## if test is true for the entire vector
  ## return a vector with true/false
  temp_yes <- yes
  ## if there is one false element, entire thing is false
  ## if all is true then it is true
  ## we have a vector , test, that TELLS us what is true and what is false
  ## how can we display either yes or no depending on test
  ## yes and no are VECTORS
  ## add into one vector either yes or no
  ## logical indexing
  ## 

  
  ## indices where test is true/false => we switch with either yes/no

  ## FOR EVERY ELEMENT!
  ## if true 
  ## yes[!true = false] <- no[!true = false]
        ## whatever index is false in x, substitute with the index in no
  ## logical indices return what is TRUE!!!!!!!
  temp_yes[!test] <- no[!test]
  temp_yes
  
}

## if(test){
## -- yes -- }
## else {
## -- no -- }


```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7b}
x <- (1:10) * pi
ifelse(x %% 1 >= 0.5, x %/% 1 + 1, x %/% 1)
my_ifelse(x %% 1 >= 0.5, x %/% 1 + 1, x %/% 1)
```

### (c)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7c}

my_sign <- function(x){
  
  my_ifelse(x > 0, x %% x+1, x %% x-1)
}



my_abs <- function(x){
  
  my_ifelse(my_sign(x) < 0, (my_sign(x) * x), x)
  
  
}

my_abs(c(-1,-2,3))
abs(c(-1,-2,3))
my_sign(c(-1,-2,3))


```
<!-- REMINDER TO USE my_ifelse() TO WRITE THE FUNCTIONS. THERE ARE OTHER WAYS TO WRITE THESE FUNCTIONS WITHOUT USING my_ifelse() BUT PLEASE FOLLOW THE DIRECTIONS. -->

## Question 8

### (a)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 8a}

merge <- function(left, right){
  merged <- numeric(0)
  while(length(left) > 0){
    
    if(length(right) > 0){
            if(left[1] < right[1]){
              merged <- c(merged, left[1])
              left <- left[-1]
              } else{
                merged <- c(merged, right[1])
                right <- right[-1]}
    } else{
      merged <- c(merged, left)
      left <- c()
    }
    
  }
  output <- c(merged,right)
  output
}



merge_sort <- function(x) {
  length_x <- length(x)
  if(length(x) > 1){
    left <- x[1:(length(x)/2)]
    right <- x[((length(x)/2)+1):length(x)]
    sorted_left <- merge_sort(left)
    sorted_right <- merge_sort(right)
    x <- merge(sorted_left, sorted_right)
  }
  x
}





```

### (b)
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 8b}
v <- merge_sort(numeric(0))
v
v <- merge_sort(7)
v

## for this case, for some reason, the 1 is ommitted?
v <- merge_sort(10:1)
v
```

## Question 9
<!-- MAKE SURE dna.RData IS IN THE SAME DIRECTORY (FOLDER) AS THIS .RMD -->
```{r Question 9 background}
load("dna.RData")
```

### (a)
<!-- CREATE THE FUNCTION AND TEST IT AS ASKED IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 9a}
locate_motif <- function(strand, motif){
  index <- numeric(0)
  count <- 0
  index <- 0
  for(i in seq_along(strand)){
    
   for(j in seq_along(motif)){
     if(strand[i] == motif[j]){
       if(count == 0)
          index <- i
       count <- count+1
     } else if(count > 0 & strand[i] != motif[j])
       count <- 0
      }
   

  }
  
  index <- index - length(motif) + 1
}
```

```{r}
seq1 <- c("A", "C", "A", "G", "T")
seq2 <- c("T", "A", "G", "T", "A")

seq3 <- c("A", "G", "T")

gimme <- locate_motif(seq1, seq3)
gimme

gimme <- locate_motif(seq2, seq3)
gimme
```

### (b)
<!-- CREATE THE FUNCTION AND TEST IT AS ASKED IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 9b}

seq1 <- c("A", "C", "A", "G", "T")
seq2 <- c("T", "A", "G", "T", "A")

extract_longest_substring <- function(strand1, strand2){
  
  longest <- 0
  temp_short <- strand1
  temp_long <- strand2
  if(length(strand2)<length(strand1)){
    temp_short <- strand2
    temp_long <- strand1
  }
  
  for(i in seq_along(temp_long)){
    
    small <- temp_short[1:i]
    
    index <- locate_motif(temp_long, small)
    
    if((index + length(small)) > longest)
    {
      longest <- (index + length(small))
    
  }
  
}
longest
}

# my logic

## go through the longest vector and if one of the elements is in common with the start of the second vector
## then we check from that INDEX and on if the elements following match the other vector.
## if not, we check the following index in the second vector
## then, we can collect the substring and as we find a longer one, we dispose of the smaller one



```



```{r}

mine <- extract_longest_substring(seq1,seq2)
mine
```
