---
title: "Stats 20, F23 -- Homework 4"
author: "Michelle Navarrete Vega -- 805676988"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

```{r}

my_t <- function(x){
  
  ## we want to get the columns and bind them as rows
  new_x <- NULL
  
  if(length(ncol(x)) != 0 ){
  my_length <- 1:ncol(x)
  
    for(i in my_length){
    x[,i]
    new_x <- rbind(new_x, x[,i])
    }
  
    }
  else {
    new_x <- cbind(x)
    }
  
  new_x
}

my_matrix <- cbind(c(1,2,3), c(4,5,6), c(7,8,9))
my_matrix


my_vec <- 1:3
my_vec
new_vec <- my_t(my_vec)
new_vec

her_vec <- cbind(1:3)
her_vec
his_vec <- my_t(her_vec)
his_vec
```




## Question 2

### Part a

```{r}

my_cor <- function(x,y, use = "everything") {
  if(use == "pairwise.complete.obs"){
      x <- x[!is.na(x)]
      y <- y[!is.na(y)]

  } 
  
  
  numerator <- (sum((x-mean(x))*(y-mean(y))))
  den_x <- sqrt(sum((x-mean(x))^2))
  den_y <- sqrt(sum((y-mean(y))^2))
  denominator <- den_x*den_y
  
  
  numerator/denominator
}

a <- 1:3
b <- 4:6
my_cor(a,b)
cor(a,b)


a1 <- c(6,2,9)
b1 <- 4:6
my_cor(a1,b1, use = "pairwise.complete.obs")
cor(a1,b1, use = "pairwise")
```

### Part b


```{r}

lin_reg <- function(x,y, use = "everything") {
  
   if(use == "pairwise.complete.obs"){
      x <- x[!is.na(x)]
      y <- y[!is.na(y)]

    } 
  
  
  b <- (cor(x,y) * (sd(y)/sd(x)))
  
  a <- mean(y) - (b*mean(x))

  
  c(a,b)
  
}

```




### Part c


```{r}


hw4_height <- c(61:64,66,68)
hw4_weight <- c(104,110,125,141,160,170)

hw4_lin <- lin_reg(hw4_height, hw4_weight)

hw4_equation <- c("y =", hw4_lin[1], "-", hw4_lin[2], "x")
hw4_equation

```

the equation is 
weight = -510.647058823529 - 10.0882352941176*height

### Part d

```{r}

linreg_mat <- function(x,y){
  
  design_matrix <- cbind(rep(1,length(x)),x)
  
  design_transpose <- my_t(design_matrix)
  
  trans_design <- design_transpose %*% design_matrix
  
  inverse <- solve(trans_design)

  output <- inverse %*% design_transpose %*% y
  
  output
  
}

```

### Part e

```{r}
linreg_mat(hw4_height,hw4_weight)
```

As in part c, the b is -510.64706 and the a is 10.08824. 

### Part f

The slope coefficient is: 10.0882352941176.

This means that for every unit value of the height, the weight is 10.0882352941176 times the height. 

It represents the change in weight for every unit value change in height.


## Question 3

### Part a

```{r}
`%m%` <- function(A,B){
  
  if(ncol(A) != nrow(B))
    stop("Matrices are not conformable.")
  
  n <- nrow(A)
  m <- ncol(B)
  output <- matrix(0, n, m)
  
    for (i in 1:n) {
    for (k in 1:m) {
      output[i, k] <- sum(A[i,] * B[,k])
    }
  }
  
  output
  
  ## should output the matrix product of A and B
}

a <- matrix(1:6, nrow = 2)
b <- matrix(7:12, ncol = 2)

a%m%b
a%*%b

```
### Part b

```{r}

X <- rbind(c(6,5,4), c(3,2,1))
Y <- cbind(c(1,2,3), c(4,5,6), c(7,8,9))

X
Y



X%m%Y

X%*%Y

```



## Question 4

### Part a

```{r}

`%^%` <- function(A,k){
  n <- nrow(A)
  result <- A
  
  if(k == 1){
    result <- A
    message(result)
  } else if(k == 0){
    result<- diag(n)
  }
  else if (k>1) {
    result <- A
  for(i in 2:k) {
    result <- result %m% A
  }
}
  result
}

a <- matrix(1:4, nrow = 2)
a%*%a
s <- 2
a%^%s

k <- 5
a%^%k
a%*%a%*%a%*%a%*%a


```
## Part b

```{r}

Z <- rbind(c(0.2, 0.7, 0.1), c(0.6, 0.2, 0.2), c(0.4, 0.1, 0.5))
Z
## k = 0
Z%^%0


## k = 5
k <- 5
Z%^%k

result <- Z
k <- 5

  for(i in 2:k) {
    result <- result %m% Z
  }
result

## k = 50
k <- 50
Z%^%k

result <- Z


  for(i in 2:k) {
    result <- result %m% Z
  }
result

## k = 500
k <- 500
Z%^%k

result <- Z


  for(i in 2:k) {
    result <- result %m% Z
  }
result

```
## Question 5

**I am going to assume that the inputs are all logical vectors and matrices**

GOAL: returns the indices of TRUE values in a logical vector or matrix

-   have a sequence of length of the vec/matrix

-   logical index, make indices false into NA

/// implement the arr.ind argument /// 
optionally return two dimensional indices of TRUE values in a logical matrix



```{r}

my_which <- function(x, arr.ind = FALSE){
  new_output <- integer(0)

  if(length(x) == 0)
    new_output <- logical(0)
  else if(length(x) == 1)
    new_output <- integer(0)
  else if (TRUE){
  ## check if it is a matrix and argument is set to true
  output <- x
  
  if(arr.ind == TRUE && length(dim(x)) == 2){
    num_rows <- 1:nrow(x)
    num_cols <- 1:ncol(x)
    rownames(output) <- c(1:nrow(x))
    colnames(output) <- c(1:ncol(x))
    ## let's try a for-loop
    new_output <- integer(0)
    for(i in num_rows){
      for(j in num_cols){
        
        if(x[i,j] == TRUE && length(new_output) == 0 )
          new_output <- c(i, j)
        else if(x[i,j] == TRUE)
          new_output <- c(new_output, c(i, j))
      }
      
    }
    
    new_output <- matrix(new_output, ncol = 2, byrow = TRUE)
    colnames(new_output) <- c("row", "col")}
                    
   else {
  
  my_len <- 1:length(x)
  
  new_output <- my_len[x]
  
  }
  }
  new_output
  
}

test <- cbind(c(TRUE, TRUE), c(FALSE,TRUE))

##my_which(test)

my_which(test, arr.ind = TRUE)

which(test, arr.ind = TRUE)

my_which(logical(1))
which(logical(1))

my_which(c(TRUE, FALSE, TRUE, TRUE, TRUE, FALSE))
which(c(TRUE, FALSE, TRUE, TRUE, TRUE, FALSE))


```


## Question 6

```{r}
load("mlb.RData")

```

### Part a


i. Find the maximum number of hits from each team

```{r}

tapply(hit, team, max)

```

ii. Find number of players on the team

-   chapter 6: how would u find out how many observations there are 'for each team'? 
```{r}

tapply(hit, team, length)
table(team)

```

iii. Using one command and no subsetting, find the number of players on each team with at least one home run

hr > 1

```{r}

tapply(hr > 0, team, sum)



```


### Part b
i. Using one command, (find the highest "batting average" for each team) among players with at least 100 at bats

players with at aleast 100 bats
find the average of those
display for each team

```{r}
tapply(hit/ab,list(team, ab>99),max)

## this was the hardest lol, not sure if it is right
## we want max, we want to compare team and ab >99
## from that list, we want hits/ab

```

ii. Using one command, find the batting average for each team.

```{r}

tapply(hit/ab,team,mean, na.rm = TRUE)

```
### Part c

Find the average number of home runs for each position in each league. Which position has the largest difference between leagues?

```{r}
tapply(hr, list(league,pos), mean)

```

Catcher: less than 1
Centerfield: 0.69
1.9
0.8
right: 4.17
shrt: 1.57



Rightfield has the largest difference between leagues. 


### Part d

Using one command, compute the median number of players for each position on any team.


```{r}


length(table(pos)) ## total

tapply(table(pos), list(1:length(table(pos))), FUN=median, na.rm = TRUE)


```

This is the only way I was able to find it. I was not able to fix the way it is presented. i think I might need to do something with the index argument.

## Question 7


### Part a

returns: a matrix of integers indicating the row number (i.e., the ijth element is equal to i) without the row() function
output of my_row(x) and row(x) should be identical for any matrix x.

```{r}

X <- cbind(c(9,3,4), c(2,5,1))

row(X)

my_row <- function(A){
  
  num_col <- 1:ncol(A)
  num_row <- 1:nrow(A)
  
  for(i in num_row)
    for(j in num_col){
      
      A[i,j] <- i }
  A
}

my_row(X)



```


### Part b

```{r}

my_col <- function(A){
  
  
  num_col <- 1:ncol(A)
  num_row <- 1:nrow(A)
  
  for(i in num_row)
    for(j in num_col){
      
      A[i,j] <- j }
  A
  
}

X <- cbind(c(9,3,4), c(2,5,1))
col(X)
my_col(X)


```


### Part c

```{r}
test_my_diag <- function(x, nrow=NULL, ncol=NULL){
  

  if(missing(x) && !is.null(nrow))
    message("return identity matrix")
  else if(length(x) == 1 && is.null(nrow) && is.null(ncol)){
    message("second case done")
    
  } else if (is.numeric(x) && (length(x)>1 || !is.null(nrow) || !is.null(ncol))){
    
    message("third case!!!!")
  }


}

m <- cbind(1:2,5:7)
  
#case 1
test_my_diag(nrow = 2)

#case 2
test_my_diag(2)

#case 3
test_my_diag(m)
  
  
  
```




returning identity matrix: 

```{r}

my_diag <- function(x, nrow=NULL, ncol=NULL){
  
  output <- integer(0)
  
  if(missing(x) && !is.null(nrow)){
    identity_matrix <- matrix(0,nrow,nrow)
    for (i in 1:nrow) {
       for (j in 1:nrow) {
       if (i == j) {
          identity_matrix[i, j] <- 1
         }
       }
    }
    output <- identity_matrix
  }else if(length(x) == 1 && is.null(nrow) && is.null(ncol)){
    
    identity_matrix <- matrix(0,x,x)
    for (i in 1:x) {
       for (j in 1:x) {
       if (i == j) {
          identity_matrix[i, j] <- 1
         }
       }
    }
        output <- identity_matrix
    
  } else if (is.numeric(x) && (length(x)>1 || !is.null(nrow) || !is.null(ncol))){
    
    my_row <- nrow(x)
    my_col <- ncol(x)
    num_diag <- x[1,1]
          for (i in 2:my_row) {
             for (j in 2:my_col) {
             if (i == j) {
                num_diag <- c(num_diag, x[i,j])
               }
             }
          }  
            output <- num_diag

    }

output
}
my_diag(nrow = 2)
diag(nrow=2)


my_diag(2)
diag(2)

m <- cbind(1:2,5:7)
my_diag(m)
diag(m)
```



## Question 8


```{r}

load("votes.RData")

```

### Part a

*Pseudocode (or outline) a function that inputs a matrix of ranked choice votes and returns a matrix of results*

(00:34):
In a ranked choice voting election, a candidate needs to earn more than half of the votes to win. 
All first choices are counted and if a candidate has a majority, then they win just like any other election. 
If not the candidate with the fewest votes is eliminated and voters who picked that candidate as a number one, will have their votes count for their next choice.
(00:54):
This process continues until a candidate earns a majority and is declared the winner.



Input a matrix. 

-   ranked choice votes

Check if any candidate earned more than half of votes.
-   If yes, they win.
  
if not
-   Eliminate least voted


Repeat process until a candidate has over half votes


Output a matrix

-   results (?)



### Part b

input is a matrix with col as names of choices
rows are votes for that choice


```{r}

tally_rcv <- function(x){
    total <- sum(apply(x, 2, sum))
    new_x <- apply(x, 2, sum)/total
    new_x <- new_x*100
    new_x
    
    ## new_x is percentages
    
    # check who has highest
    
    ## if not over 50, then eliminate lowest and do again
        ## find total again and repeat the process.
  
}


tally_rcv(votes)

```

## part c

incomplete