---
title: "Stats 20, F23 -- Homework 7"
author: "Michelle Navarrete Vega -- 805676988"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

### Part a

```{r}

births <- read.csv("births.csv", stringsAsFactors = TRUE)
is.factor(births$Habit)

```

### Part b

```{r}

table(births$Habit)

names(table(births$Habit)[1])

levels(births$Habit)


```


There are 6 observations in the category where there is no information regarding the smoking habits of the mother. The name of this category is "" blank.


### Part c

```{r}

plot(births$Habit, col = "lightgreen")

```

### Part d

```{r}

known_habit_data <- births[births$Habit %in% c("NonSmoker", "Smoker"), ]

known_habit_data$Habit <- droplevels(known_habit_data$Habit)

levels(known_habit_data$Habit)

```


### Part e

```{r}

boxplot(known_habit_data$weight~known_habit_data$Habit)

```




## Question 2

```{r}

with(births, hist(weight[Habit == "Smoker"],
            freq = FALSE,density = 15, col = "red", 
            xlab = "Weight (in ounces)", main = "Histogram of Weight by Mother's Smoking Habit",
            xlim = range(weight), ylim = c(0,0.03)
                  ))


with(births, hist(weight[Habit == "NonSmoker"],
            freq = FALSE, density = 30, col = "darkgreen", 
            breaks = 20, add = TRUE
                  ))

legend("topleft", c("Smoker", "NonSmoker", "Smoker Density Estimate", "NonSmoker Density Estimate", "Smoker Median", "NonSmoker Median"),
       density = c(15, 30, NA,NA,NA,NA),
       fill = c("red", "darkgreen",0,0,0,0), border= c(1,1,NA,NA,NA,NA), 
       inset = 0.005, col = c(NA,NA, "red", "darkgreen", "red", "darkgreen"), lty = c(NA,NA, 1,1,2,2), lwd = c(NA,NA,1.5,1.5,1,1))

lines(density(births$weight[births$Habit == "Smoker"]), lwd = 1.5, col = "red")
lines(density(births$weight[births$Habit == "NonSmoker"]), lwd = 1.5, col = "darkgreen")

abline(v = summary(births$weight[births$Habit == "Smoker"])[3], lty = 2, col = "red")
abline(v = summary(births$weight[births$Habit == "NonSmoker"])[3], lty = 2, col = "darkgreen")




```


## Question 3

### Part a

```{r chunk_name, cache = TRUE}
library(ggplot2)
pairs(diamonds[, c("carat", "depth", "table", "price")], lower.panel = NULL, cex = 0.1)


```
carat and price seem to have the strongest relationship

### Part b

```{r}

plot(price ~ carat, data = diamonds, cex = 0.3, pch = 4, col = clarity, main = "Carat vs. Price")
legend("topright", 
       legend = levels(diamonds$clarity), 
       col = 1:length(diamonds$clarity), 
       pch = 4,
       title = "Clarity",  inset = 0.005)

## class(diamonds$clarity)
## clarity is an ordered factor. 

```
The colors are assigned based on my R's internal color palette, which can be customized. Each color is assigned a number and since there are 8 levels in clarity, each level will be of the color of that number. Ordered factors have a natural ordering, and the default color assignment reflects this order.


### Part c

```{r}
custom_colors <- c("coral", "blue", "deeppink", "darkmagenta", "chartreuse4", "gold", "darkred", "cornflowerblue")

# Map clarity levels to custom colors
clarity_colors <- custom_colors[as.numeric(diamonds$clarity)]

plot(price ~ carat, data = diamonds, cex = 0.3, pch = 4, col = clarity_colors, main = "Carat vs. Price")

legend("topright", 
       legend = levels(diamonds$clarity), 
       col = custom_colors, 
       pch = 4,
       title = "Clarity", inset = 0.02)

## class(diamonds$clarity)
## clarity is an ordered factor. 

```


### Part d

The higher the price, the diamonds more likely to be of "internally flawless". Lower clarity grades, such as Included 1 (I1), are generally associated with lower-priced diamonds. Diamonds with more visible inclusions are considered less valuable. Diamonds with higher clarity grades tend to be more expensive. Heavier diamonds are more rare. We can also observe a trend where diamonds have larger carat weights but possibly lower prices due to the lower clarity grade. I've learned how "Internally Flawless" (IF) are considered very high-quality diamonds with no visible inclusions under 10x magnification. These diamonds are rarer and, therefore, tend to command higher prices. 



## Question 4

### Part a

```{r}

mean_cut_color <- matrix(
  tapply(diamonds$price, list(diamonds$color, diamonds$cut), mean),
  nrow = length(levels(diamonds$color))
)

```


### Part b

```{r}

# Create a matrix of mean prices for each color-cut combination
mean_cut_color <- matrix(
  tapply(diamonds$price, list(diamonds$color, diamonds$cut), mean),
  nrow = length(levels(diamonds$color))
)

rownames(mean_cut_color) <- levels(diamonds$color)
colnames(mean_cut_color) <- levels(diamonds$cut)

matplot(1:length(levels(diamonds$color)), mean_cut_color, type = "l", lty = 1:3, col = 1:6, lwd = 2, xaxt = "n", xlab = "Color", ylab = "Mean Price")

axis(1, at = 1:length(levels(diamonds$color)), labels = levels(diamonds$color))

legend("topright", legend = levels(diamonds$cut), col = 1:6, lty = 1:6, lwd = 2, title = "Cut")


```


### Part c

In ordered rank, color levels closer to J are of higher price. For cut, based on the ordered rank, it varies. Premium costs the highest, ideal costs the least. In increasing order: ideal, good, very good, fair, premium. 

```{r}
mean_cut_color
summary(mean_cut_color)
```

Premium has the highest mean price, while Ideal has the lowest mean price. 

